Vagueness, Noise, and Signaling

### 2014-11-05 paper/paper_short.tex
cover letter

% \linespread{1.5}

@@

\begin{abstract}
  Signaling games are popular models for studying the evolution of
  meaning, but typical approaches do not allow for vagueness as a
  feature of successful signaling.  Complementing recent like-minded
  models, we introduce the replicator diffusion dynamic as a special
  case of the replicator mutator dynamic that combines evolutionary
  pressure on optimal signal use with stochastic noise in the
  processing of similar stimuli. Applying this new dynamic to a
  generalization of Lewis' signaling games, we show that stochastic
  imprecision leads to vague, yet communicative efficient signal use,
  and, moreover, also unifies evolutionary outcomes and helps avoid
  suboptimal categorization.
\end{abstract}

Many of our concepts and words are vague. A vague category knows clear
cases that fall under it, clear cases that do not, and also so-called
borderline cases. Borderline cases do not clearly apply, nor do they
clearly not apply, and there may be differences between borderline
cases in terms of how well they represent the category in
question. Vagueness does not seem to dramatically affect the success
of everyday communication, but it is troublesome for some of the most
prominent theories of language and meaning. This is especially the
case for the logico-semantic tradition of Frege, Russell and the young
Wittgenstein which is challenged by the paradoxes vagueness gives rise
to.

@@

The replicator diffusion dynamic was first introduced by \citet{Correia2013:The-Bivalent-Tr}
as a noise-perturbed variant of the replicator dynamic (\rd) in
behavioral strategies. Fix a sim-max game with
$\States = \Acts$ and a confusion matrix $C : \States \times
\States$. $C$ is a row-stochastic matrix whose elements $C_{ij}$ give
the probability that $\state_i$ is realized as $\state_j$. The
confusability of states affects senders and receivers alike, but in
slightly different ways (see
Figure~\ref{fig:noise-perturbation-of-strategies}). For the sender,
$C_{ij}$ is the probability that the actual state \mystate{i} is
perceived as \mystate{j}. For the receiver, $C_{ij}$ is the
probability that \mystate{j} is the interpretation that is actually
formed when \mystate{i} is the intended interpretation.

@@

The discrete-time replicator diffusion dynamic takes the replicator
dynamic as basic, but factors in the confusion of states at each
update step in a sequential update:
\begin{align*}
  \RDD(\Sstrat) & = \Diff_C(\RD(\Sstrat)) &   \RDD(\Rstrat) & = \Diff_C(\RD(\Rstrat)) \,.
\end{align*}
This is equivalent to the following, perhaps more transparent
formulation, given by \citet{Correia2013:The-Bivalent-Tr}:
\begin{align*}
  \RDD(\Sstrat)(\messg \probbar \state_i) & = \sum_{j} C_{ij} \cdot
  \frac{\Sstrat(\messg \probbar \state_j) \cdot
    \EU(\messg, \state_j,\Rstrat)}
  {\Phi(\state_j,\Sstrat, \Rstrat)} \\
    \RDD(\Rstrat)(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
  \frac{\Rstrat(\state_j \probbar \messg) \cdot
    \EU(\state_j, \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
\end{align*}

>>>>

\doublespacing

@@

\begin{abstract}
  Signaling games are popular models for studying the evolution of
  meaning, but typical approaches do not incorporate vagueness as a
  feature of successful signaling.  Complementing recent like-minded
  models, we introduce the replicator diffusion dynamic as a special
  case of the replicator mutator dynamic that combines evolutionary
  pressure on optimal signal use with stochastic noise in the
  processing of similar stimuli. Applying this new dynamic to a
  generalization of Lewis' signaling games, we show that stochastic
  imprecision leads to vague, yet communicative efficient signal use,
  and, moreover, unifies evolutionary outcomes and helps avoid
  suboptimal categorization.
\end{abstract}

Many concepts and expressions are vague. A vague category knows clear
cases that fall under it, clear cases that do not, and also so-called
borderline cases. Borderline cases do not clearly apply, nor do they
clearly not apply, and there may be differences between borderline
cases in terms of how well they represent the category in
question. Vagueness does not seem to dramatically affect the success
of everyday communication, but it is troublesome for some of the most
prominent theories of language and meaning. This is especially so for the logico-semantic tradition of Frege, Russell and the young
Wittgenstein which is challenged by the paradoxes vagueness gives rise
to.

@@

The replicator diffusion dynamic is a noise-perturbed variant of the
replicator dynamic (\rd) in behavioral strategies. Fix a sim-max game
with $\States = \Acts$ and a confusion matrix $C : \States \times
\States$. $C$ is a row-stochastic matrix whose elements $C_{ij}$ give
the probability that $\state_i$ is realized as $\state_j$. The
confusability of states affects senders and receivers alike, but in
slightly different ways (see
Figure~\ref{fig:noise-perturbation-of-strategies}). For the sender,
$C_{ij}$ is the probability that the actual state \mystate{i} is
perceived as \mystate{j}. For the receiver, $C_{ij}$ is the
probability that \mystate{j} is the interpretation that is actually
formed when \mystate{i} is the intended interpretation.

@@

The discrete-time replicator diffusion dynamic takes the replicator
dynamic as basic, but factors in the confusion of states at each
update step in a sequential update:
\begin{align*}
  \RDD(\Sstrat) & = \Diff_C(\RD(\Sstrat)) &   \RDD(\Rstrat) & = \Diff_C(\RD(\Rstrat)) \,.
\end{align*}
This is equivalent to the following, perhaps more transparent
formulation:
\begin{align*}
  \RDD(\Sstrat)(\messg \probbar \state_i) & = \sum_{j} C_{ij} \cdot
  \frac{\Sstrat(\messg \probbar \state_j) \cdot
    \EU(\messg, \state_j,\Rstrat)}
  {\Phi(\state_j,\Sstrat, \Rstrat)} \\
    \RDD(\Rstrat)(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
  \frac{\Rstrat(\state_j \probbar \messg) \cdot
    \EU(\state_j, \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
\end{align*}
